---
title: "TBANLT 560 Project 2- FINAL!!!"
author: "Lindsey Grays"
date: "03/18/2021"
output:
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/lindseyg/OneDrive - Bill & Melinda Gates Foundation/TBANLT 560"))
```
 
```{r}
library(mlbench)
library(e1071)
library(klaR)
library(nnet)
library(MASS)
library(rpart)
library(randomForest)
library(ggplot2)
library(caret)
library(nnet)
library(caretEnsemble)
```
## LOAD DATA IN
```{r}
data("BreastCancer")
```

### GET RID OF ID COLUMN AND NAS TO AVOID ISSUES WHEN MODELING
### TAKE A LOOK AT THE SUMMARY STATS AND STRUCTURE OF THE DATA
```{r}
BreastCancer <- na.omit(BreastCancer) 
BreastCancer$Id <- NULL
summary(BreastCancer)
str(BreastCancer)
```

### LOOK AT THE DISTRIBUTION OF BENIGN VS MALIGNANT
### THANKFULLY IT APPEARS THAT MOST TUMORS ARE BENIGN
```{r}
ggplot(data = BreastCancer, aes(x = Class, fill = Class)) +
  geom_bar() + labs(x= "Type of Breast Tumor Present", y = "Total Number of Patients")
```
```{r}
round(prop.table(table(BreastCancer$Class)),3)
```


### DATA EXPLORATION
### PLOT THE VARIABLES
```{r}
plot(BreastCancer)
```

### GRAPH ALL INDEPENDENT VARIABLES AGAINST THE DEPENDENT ONE (CLASS)
```{r}
ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Cl.thickness),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("CI Thickness")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Cell.size),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("Cell Size")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Cell.shape),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("Cell Shape")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Marg.adhesion),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("Marg Adhesion")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Bare.nuclei),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("Bare Nuclei")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Bl.cromatin),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("Bl Cromatin")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Normal.nucleoli),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("Normal Nucleoli")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Mitoses),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("Mitoses")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

ggplot()+
  geom_bar(data = BreastCancer, aes(x=factor(BreastCancer$Epith.c.size),fill=factor(BreastCancer$Class)),
           position = "stack")+
  scale_x_discrete("Epith.c.size")+
  scale_y_continuous("Number of Patients")+
  guides(fill=guide_legend(title="Class"))+
  scale_fill_manual(values=c("blue","red"))

```

### TAKE A LOOK AT WHAT VARIABLES ARE MOST CORRELATED
### IT APPEARS THAT THE MOST CORRELATED WITH CLASS ARE CELL.SIZE, CELL.SHAPE, AND BARE.NUCLEI AT .82; CI THICKNESS, MARG.ADHESION, BL. CROMATIN, AND NORMAL NUCLEOLI ARE ALSO HIGHLY CORRELATED WITH CLASS. MITOSIS IS THE LEAST CORRELATED WITH CLASS

```{r}
library(Hmisc)
data("BreastCancer")
BreastCancer <- na.omit(BreastCancer) 
BreastCancer$Id <- NULL
BreastCancer$Class<-as.numeric(BreastCancer$Class)
BreastCancer$Cl.thickness<-as.numeric(BreastCancer$Cl.thickness)
BreastCancer$Cell.size<-as.numeric(BreastCancer$Cell.size)
BreastCancer$Cell.shape<-as.numeric(BreastCancer$Cell.shape)
BreastCancer$Marg.adhesion<-as.numeric(BreastCancer$Marg.adhesion)
BreastCancer$Epith.c.size<-as.numeric(BreastCancer$Epith.c.size)
BreastCancer$Bare.nuclei<-as.numeric(BreastCancer$Bare.nuclei)
BreastCancer$Bl.cromatin<-as.numeric(BreastCancer$Bl.cromatin)
BreastCancer$Normal.nucleoli<-as.numeric(BreastCancer$Normal.nucleoli)
BreastCancer$Mitoses<-as.numeric(BreastCancer$Mitoses)
M <-cor(BreastCancer)
M
```

### SPLIT DATA INTO TRAINING (80%) AND EVALUATION SET (20%)
```{r}
data("BreastCancer")
BreastCancer <- na.omit(BreastCancer) 
BreastCancer$Id <- NULL
set.seed(2)
partition <- createDataPartition(BreastCancer$Class, p =0.8, list = FALSE)
BreastCancer.train<- BreastCancer[partition, ]
cat("the dimension of the training dataset is (", dim(BreastCancer.train), ")")
```
```{r}
BreastCancer.test <- BreastCancer[-partition,]
cat("the dimension of the test dataset is (", dim(BreastCancer.test), ")")
#ALTERNATIVE CODE:
#ind <- sample(2, nrow(BreastCancer), replace = TRUE, prob=c(0.8, 0.2))
```

############################
MODELING
############################

# SVM
```{r}
svmRadial <- svm(Class ~ ., BreastCancer.train)
mysvm.pred <- predict(svmRadial, BreastCancer.train)
table(mysvm.pred,BreastCancer.train$Class)
```


# NB
```{r}
naive_bayes <- NaiveBayes(Class ~ ., BreastCancer.train)
mynb.pred <- predict(naive_bayes,BreastCancer.train)
table(mynb.pred$class,BreastCancer.train$Class)
```

# NNET
```{r}
nnet <- nnet(Class ~ ., BreastCancer.train, size=1)
mynnet.pred <- predict(nnet,BreastCancer.train,type="class")
table(mynnet.pred,BreastCancer.train$Class)
```

# Decision trees
```{r}
library(rpart)
rpart <- rpart(Class ~ ., BreastCancer.train)
plot(rpart); text(rpart) # in "iris_tree.ps"
summary(rpart)
mytree.pred <- predict(rpart,BreastCancer.train,type="class")
table(mytree.pred,BreastCancer.train$Class)
```
# Leave-1-Out Cross Validation (LOOCV)
```{r}
ans <- numeric(length(BreastCancer.train[,1]))
for (i in 1:length(BreastCancer.train[,1])) {
  mytree <- rpart(Class ~ ., BreastCancer.train[-i,])
  mytree.pred <- predict(mytree,BreastCancer.train[i,],type="class")
  ans[i] <- mytree.pred
}
ans <- factor(ans,labels=levels(BreastCancer.train$Class))
table(ans,BreastCancer.train$Class)
```


# Quadratic Discriminant Analysis
```{r}
BreastCancer.train$Class<-as.numeric(BreastCancer.train$Class)
BreastCancer.train$Cl.thickness<-as.numeric(BreastCancer.train$Cl.thickness)
BreastCancer.train$Cell.size<-as.numeric(BreastCancer.train$Cell.size)
BreastCancer.train$Cell.shape<-as.numeric(BreastCancer.train$Cell.shape)
BreastCancer.train$Marg.adhesion<-as.numeric(BreastCancer.train$Marg.adhesion)
BreastCancer.train$Epith.c.size<-as.numeric(BreastCancer.train$Epith.c.size)
BreastCancer.train$Bare.nuclei<-as.numeric(BreastCancer.train$Bare.nuclei)
BreastCancer.train$Bl.cromatin<-as.numeric(BreastCancer.train$Bl.cromatin)
BreastCancer.train$Normal.nucleoli<-as.numeric(BreastCancer.train$Normal.nucleoli)
BreastCancer.train$Mitoses<-as.numeric(BreastCancer.train$Mitoses)
qda <- qda(Class ~ ., BreastCancer.train)
myqda.pred <- predict(qda, BreastCancer.train)
table(myqda.pred$class,BreastCancer.train$Class)
```

# Regularised Discriminant Analysis
```{r}
data("BreastCancer")
BreastCancer <- na.omit(BreastCancer) 
BreastCancer$Id <- NULL
set.seed(2)
partition <- createDataPartition(BreastCancer$Class, p =0.8, list = FALSE)
BreastCancer.train<- BreastCancer[partition, ]
BreastCancer.test <- BreastCancer[-partition,]
```
```{r}
binda <- rda(Class ~ ., BreastCancer.train)
myrda.pred <- predict(binda, BreastCancer.train)
table(myrda.pred$class,BreastCancer.train$Class)
```

# Random Forests
```{r}
rf <- randomForest(Class ~ .,BreastCancer.train)
myrf.pred <- predict(rf, BreastCancer.train)
table(myrf.pred, BreastCancer.train$Class)
```
## WOW!

###############
ENSEMBLE
##############
#stack

```{r}
set.seed(100)

control_stacking <- trainControl(method="repeatedcv", number=5, repeats=2, savePredictions=TRUE, classProbs=TRUE)

algorithms_to_use <- c('rpart', 'svmRadial', 'rf', 'naive_bayes', 'nnet')

stacked_models <- caretList(Class ~., data=BreastCancer.train, trControl=control_stacking, methodList=algorithms_to_use)

stacking_results <- resamples(stacked_models)

summary(stacking_results)
```
```{r}
# stack using glm
stackControl <- trainControl(method="repeatedcv", number=5, repeats=3, savePredictions=TRUE, classProbs=TRUE)

set.seed(100)

glm_stack <- caretStack(stacked_models, method="glm", metric="Accuracy", trControl=stackControl)

print(glm_stack)
```



